---
title: "HW1-Summary"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{HW1-Summary}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(bis557)
```


### This is a grader-friendly checklist for HW#1.

### 1. 
Create an R package called 'bis557'.

### 2. 
Check 'linear-model.R' in 'R' folder, which includes the function and documentation.

### 3.
Check 'lm_patho.rda' in 'data' folder.$\\$

Check 'DATASET.R' in 'data-raw' folder, which contains the code to read in raw dataset.$\\$

Check 'lm_patho.R' in 'R' folder, which includes a brief documentation of the data.$\\$

### 4.
Check 'test-linear-model.r' in folder 'tests' -> 'testthat'.

### 5.
Check 'gradient-descent.R' in 'R' folder, which includes the function and documentation.

### 6.
Check 'test-gradient-descent.R' in folder 'tests' -> 'testthat'.

### 7 & 8.
Check badges on Github page.

